Student Log-Data from a Randomized Evaluation of Educational Technology: A Causal Case Study

### 2019-10-11 paper.Rnw
main model drops missing log data

Over the course of the effectiveness study, Carnegie Learning gathered
log data for each worked problem.
The resulting dataset lists the problem name, section, and unit of
each problem, the numbers of hints requested and errors committed, and time-stamps.
Log data were missing for some students, either because the log files were not retrievable, or because of an imperfect ability to link log data to other student records.
Treatment schools with mastery data missing
for 90\% or more students were omitted from the analysis, along with
their entire randomization block.
Of the remaining 2,390 students, 87\% (2,091) had log
data.
The students with missing log data were dropped from the observational
study and mediation analysis, but included in the principal
stratification study.
Problem log data for sections that were not part of the
standard CTA1 Algebra I curriculum and problems worked by fewer than 100
students were omitted from the
dataset.

All told, the main analysis included $n=$5308 students, 2390 of whom were assigned to the CTA1 condition and 2918 of whom were assigned to control.
The students were nested within 116 teachers, in 43 schools across five states.

>>>>

Over the course of the effectiveness study, Carnegie Learning gathered
log data for each worked problem.
The resulting dataset lists the problem name, section, and unit of
each problem, the numbers of hints requested and errors committed, and time-stamps.
Log data were missing for some students, either because the log files were not retrievable, or because of an imperfect ability to link log data to other student records.
Treatment schools with mastery data missing
for 90\% or more students were omitted from the analysis, along with
their entire randomization block.
Of the remaining \Sexpr{Nt} students, \Sexpr{round(obsUse/Nt*100)}\% (\Sexpr{prettyNum(obsUse,',')}) had log
data.
For the sake of simplicity, the treatment students without log data
were dropped from the study.\footnote{Including these students in a
  principal stratification model is straightforward \citep{aoas}; see
  the online supplement for results from a model including them, which
  were broadly similar to the results presented here. Including
  subjects with missing log data in a mediation or observational study
  design can be more problematic \citep[see, e.g.][]{li2017identifiability}.}
Problem log data for sections that were not part of the
standard CTA1 Algebra I curriculum and problems worked by fewer than 100
students were omitted from the
dataset.

All told, the main analysis included $n=$\Sexpr{prettyNum(sdat$nstud,',')} students, \Sexpr{prettyNum(sum(sdat$Z),',')} of whom were assigned to the CTA1 condition and \Sexpr{prettyNum(with(sdat,nstud-sum(Z)),',')} of whom were assigned to control.
The students were nested within \Sexpr{sdat[['nteacher']]} teachers, in \Sexpr{sdat[['nschool']]} schools across five states.
